---
title: "penguins_ANOVA"
author: "Nelson Gould"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read-in penguins.csv data

```{r, messages = FALSE}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```


Our questions

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, Adelie)?

2. Do some exploring 

New functions: with + table for easy counting over multiple factors 

Update species name using case_when

```{r}

counts <- with(penguins, table(species, sex))

counts

penguins2 <- penguins %>% 
  mutate(common_name = 
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
           )) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")

penguins2

```

- Exploratory graphs
- Summary statistics
- Levene's tests for equal variances
- one-way ANOVA
- post-hoc testing

```{r}

hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) 

hists

#Making QQ-plots

qq_pen <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq() +
  facet_wrap(~ common_name, scales = "free")

qq_pen

# Levene's test for equal variances (> 2 groups...if you have two groups, use the F-test for equal variances)

# We'll use the function leveneTest from the 'car' package

# H0: Variances are equal 
# HA: Variances are unequal

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)
penguin_levene

# We retain the null hypothesis of equal variances (p = 0.06)

sum_table_pen <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    var_flip = var(flipper_length)
  )
sum_table_pen


```

3. One-way ANOVA

Single factor: common_name (species)
Number of levels in the factor: 3 (Chinstrap, Gentoo, Adelie)
Random variable: flipper_length (mm)

Is there a significant difference in mean flipper length (mm) for femaile Chinstrap, Gentoo, and Adelie penguins?

```{r}

#H0: Mean flipper lengths across all species are equal 
#HA: At least two means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

# If we draw random samples from a population there is an almost 0% chance that those sample means equal the sample means we found
# At least two samples were taken from populations with different means.

#Which ones are different? All three different from each other? Or something else?

# Post-hoc testing using Tukey's HSD

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph
```

Mean flipper lengths for Adelie, Gentoo, and Chinstrap penguins are all significantly different (F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).









